---
title: "Comparison of model performance"
author: "Marieke Vinkenoog"
date: "26/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyr)
library(ggstance)
library(stringr)
library(ggh4x)
```

# Functions to load data 
```{r}
load_files <- function(folder, target_name) {
  info <- str_split(folder, '-', simplify=TRUE)       # country, size
  file <- paste0(path, folder, target_name)
  data <- read.csv(file) %>%
    mutate(country = info[1],
           size = info[2])
  return(data)
}
```

# Loading files 
These are all runs on mode 'final'.

```{r}
path <- '../../results/v0.32_results/'

folders <- list.dirs(path, full.names=FALSE, recursive=FALSE)
folders_10k <- grep('10k', folders, value=TRUE)

# sizes.csv is needed to calculate deferral rate in train/validate sets
datalist <- lapply(folders_10k, load_files, target_name='/sizes.csv')
data_sizes <- bind_rows(datalist) %>%
  separate(Id, into=c(NA, 'sex'), sep='-') %>%
  distinct() %>%
  mutate(deferral_rate = Deferred.last.donations / Donors,
         country = case_when(country=='BE' ~ 'Belgium',
                             country=='NL' ~ 'Netherlands',
                             country=='FL' ~ 'Finland',
                             country=='SA' ~ 'South Africa',
                             country=='AU' ~ 'Australia')) %>%
  select(country, sex, Dataset, deferral_rate) %>%
  pivot_wider(names_from=Dataset, names_prefix='defrate_', values_from=deferral_rate)

# summary.csv contains the performance measures
datalist <- lapply(folders_10k, load_files, target_name='/summary.csv')
data_aupr <- bind_rows(datalist) %>%
  select(country, Pretty, Sex, AUPR.value, AUPR.low, AUPR.high) %>%
  mutate(country = case_when(country=='BE' ~ 'Belgium',
                             country=='NL' ~ 'Netherlands',
                             country=='FL' ~ 'Finland',
                             country=='SA' ~ 'South Africa',
                             country=='AU' ~ 'Australia'),
         country = as.factor(country),
         model = factor(str_to_title(Pretty), 
                           levels=c('Baseline',
                                    'Random Forest',
                                    'Support Vector Machine',
                                    'Linear Mixed Model',
                                    'Dynamic Linear Mixed Model')),
         sex = as.factor(Sex)) %>%
  select(-Pretty, -Sex) %>%
  merge(data_sizes, by=c('country', 'sex'), all.x=TRUE) %>%
  mutate(AUPR.value.adj = (AUPR.value - defrate_validate),
         AUPR.low.adj   = AUPR.low   - defrate_validate,
         AUPR.high.adj  = AUPR.high  - defrate_validate)
```

# Plotting AUPR per model

```{r}
data_aupr %>% 
  ggplot(aes(x=AUPR.value.adj, xmin=AUPR.low.adj, xmax=AUPR.high.adj, y=country)) +
    facet_grid(rows=vars(model), cols=vars(sex), labeller=labeller(model=label_wrap_gen(10),
                                                                   sex=str_to_title)) +
    ggstance::geom_pointrangeh(aes(colour=country, shape=country)) +
    scale_shape_manual(values=c(18, 15, 8, 17, 16)) +
    scale_y_discrete(limits=rev) +
    xlab('Area Under Precision-Recall Curve, adjusted for deferral rate') +
    ylab('') +
    theme_bw() +
    theme(legend.position='none',
          strip.text.y.right=element_text(angle=0))

ggsave('../plots_for_manuscript/AUPR_comparison.png', width=5, height=5)

data_aupr %>% 
  ggplot(aes(x=AUPR.value.adj, xmin=AUPR.low.adj, xmax=AUPR.high.adj, y=model)) +
    facet_grid(rows=vars(country), cols=vars(sex), labeller=labeller(sex=str_to_title)) +
    ggstance::geom_pointrangeh(aes(colour=country, shape=country)) +
    scale_shape_manual(values=c(18, 15, 8, 17, 16)) +
    scale_y_discrete(limits=rev) +
    xlab('Area Under Precision-Recall Curve, \nadjusted for deferral rate') +
    ylab('') +
    theme_bw() +
    theme(legend.position='none',
          strip.text.y.right=element_text(angle=0))

ggsave('../plots_for_manuscript/AUPR_comparison_bycountry.png', width=5, height=5)

```
# Plotting deferral rate vs AUPR 

```{r}
data_aupr %>%
  ggplot(aes(x=defrate_validate, y=AUPR.value.adj, color=country, shape=country)) +
  facet_grid(rows=vars(model), cols=vars(sex), labeller=labeller(model=label_wrap_gen(10),
                                                                   sex=str_to_title),
             scales='free_x') +
  geom_point(size=2) +
  scale_shape_manual(values=c(18, 15, 8, 17, 16)) +
  xlab('Deferral rate in validation data') +
  ylab('Area Under Precision-Recall Curve, \nadjusted for deferral rate') +
  theme_bw() +
    theme(legend.position='bottom',
          strip.text.y.right=element_text(angle=0)) +
  guides(shape=guide_legend(nrow=2, byrow=TRUE))

ggsave('../plots_for_manuscript/defrate_vs_AUPR.png', width=5, height=6)
```


# Deferral rate vs AUPR in multiple Belgian runs
Preprocessed data with different proportions of deferrals removed, to see if
AUPR drops as deferral rate gets smaller.

```{r}
load_files_BE <- function(folder, target_name) {
  info <- str_split(folder, '-', simplify=TRUE)       # country, size
  file <- paste0(path, folder, target_name)
  data <- read.csv(file) %>%
    mutate(country = info[1],
           size = info[2],
           defprop = info[3])
  return(data)
}

path <- '../../results/BE extra runs deferral rate/'

folders <- list.dirs(path, full.names=FALSE, recursive=FALSE)

datalist <- lapply(folders, load_files_BE, target_name='/sizes.csv')
data_sizes <- bind_rows(datalist) %>%
  separate(Id, into=c(NA, 'sex'), sep='-') %>%
  distinct() %>%
  mutate(deferral_rate = Deferred.last.donations / Donors,
         country = case_when(country=='BE' ~ 'Belgium',
                             country=='NL' ~ 'Netherlands',
                             country=='FL' ~ 'Finland',
                             country=='SA' ~ 'South Africa')) %>%
  select(country, sex, Dataset, deferral_rate, defprop) %>%
  pivot_wider(names_from=Dataset, names_prefix='defrate_', values_from=deferral_rate)

# summary.csv contains the performance measures
datalist <- lapply(folders, load_files_BE, target_name='/summary.csv')
data_aupr <- bind_rows(datalist) %>%
  select(country, Pretty, Sex, AUPR.value, AUPR.low, AUPR.high, defprop) %>%
  mutate(country = case_when(country=='BE' ~ 'Belgium',
                             country=='NL' ~ 'Netherlands',
                             country=='FL' ~ 'Finland',
                             country=='SA' ~ 'South Africa'),
         country = as.factor(country),
         model = factor(str_to_title(Pretty), 
                           levels=c('Baseline',
                                    'Random Forest',
                                    'Support Vector Machine',
                                    'Linear Mixed Model',
                                    'Dynamic Linear Mixed Model')),
         sex = as.factor(Sex)) %>%
  select(-Pretty, -Sex) %>%
  merge(data_sizes, by=c('country', 'sex', 'defprop'), all.x=TRUE) %>%
  mutate(AUPR.value.adj = (AUPR.value - defrate_validate),
         AUPR.low.adj   = AUPR.low   - defrate_validate,
         AUPR.high.adj  = AUPR.high  - defrate_validate,
         defprop = as.factor(as.numeric(defprop)))

data_aupr %>%
  ggplot(aes(x=defrate_validate, y=AUPR.value.adj, color=defprop, shape=defprop)) +
  facet_grid(rows=vars(model), cols=vars(sex), labeller=labeller(model=label_wrap_gen(10),
                                                                   sex=str_to_title),
             scales='free_x') +
  geom_point(size=2) +
  xlab('Deferral rate in validation data') +
  ylab('Area Under Precision-Recall Curve, \nadjusted for deferral rate') +
  guides(shape=guide_legend(title='Percentage of original deferral rate'),
         color=guide_legend(title='Percentage of original deferral rate')) +
  theme_bw() +
    theme(legend.position='bottom',
          strip.text.y.right=element_text(angle=0))

ggsave('../plots_for_manuscript/defrate_vs_AUPR_BE.png', width=5, height=5)
```



# Loading results on full datasets

```{r}
folders_full <- grep('full|50k', folders, value=TRUE)

# sizes.csv is needed to calculate deferral rate in train/validate sets
datalist <- lapply(folders_full, load_files, target_name='/sizes.csv')
data_sizes_full <- bind_rows(datalist) %>%
  separate(Id, into=c(NA, 'sex'), sep='-') %>%
  distinct() %>%
  mutate(deferral_rate = Deferred.last.donations / Donors,
         country = case_when(country=='BE' ~ 'Belgium',
                             country=='NL' ~ 'Netherlands',
                             country=='FL' ~ 'Finland',
                             country=='SA' ~ 'South Africa')) %>%
  select(country, sex, Dataset, deferral_rate) %>%
  pivot_wider(names_from=Dataset, names_prefix='defrate_', values_from=deferral_rate)

# summary.csv contains the performance measures
datalist <- lapply(folders_full, load_files, target_name='/summary.csv')
data_aupr_full <- bind_rows(datalist) %>%
  select(country, Pretty, Sex, size, AUPR.value, AUPR.low, AUPR.high) %>%
  mutate(country = case_when(country=='BE' ~ 'Belgium',
                             country=='NL' ~ 'Netherlands',
                             country=='FL' ~ 'Finland',
                             country=='SA' ~ 'South Africa'),
         country = as.factor(country),
         model = factor(str_to_title(Pretty), 
                           levels=c('Baseline',
                                    'Random Forest',
                                    'Support Vector Machine',
                                    'Linear Mixed Model',
                                    'Dynamic Linear Mixed Model')),
         sex = as.factor(Sex)) %>%
  select(-Pretty, -Sex) %>%
  merge(data_sizes_full, by=c('country', 'sex'), all.x=TRUE) %>%
  mutate(AUPR.value.adj = (AUPR.value - defrate_validate),
         AUPR.low.adj   = AUPR.low   - defrate_validate,
         AUPR.high.adj  = AUPR.high  - defrate_validate,
         size = 'full')

data_aupr_10k <- data_aupr %>%
  mutate(size = '10k') 

data_aupr <- data_aupr_10k %>%
  bind_rows(data_aupr_full)
```


# Comparing model performance in 10k sample and full data

```{r}
data_aupr %>%
  filter(country != 'Belgium') %>%
  ggplot(aes(x=AUPR.value.adj, xmin=AUPR.low.adj, xmax=AUPR.high.adj, y=country)) +
  facet_grid(rows=vars(model), cols=vars(sex), labeller=labeller(model=label_wrap_gen(10),
                                                                   sex=str_to_title)) +
  ggstance::geom_pointrangeh(aes(colour=country, shape=interaction(country,size))) +
  scale_shape_manual(values=c(15, 16, 17, 0, 1, 2)) +
    scale_y_discrete(limits=rev) +
    xlab('Area Under Precision-Recall Curve, adjusted for deferral rate') +
    ylab('') +
    theme_bw() +
    theme(legend.position='bottom',
          strip.text.y.right=element_text(angle=0))

  
  
  
  

```

# Score on full data relative to 10k data

```{r}
data_rel <- data %>%
  group_by(Group, measure) %>%
  mutate(value_rel2 = value - value[Size == '10k'],
         low_rel2 = low - low[Size == '10k'],
         high_rel2 = high - high[Size == '10k']) %>%
  ungroup()



data_rel %>% 
  filter(Country %in% c('Finland', 'Netherlands'),
         Size == 'full') %>%
  ggplot(aes(x=value_rel2, y=Group)) +
    facet_grid(rows=vars(Pretty), cols=vars(measure)) +
    geom_point(aes(colour=Sex, shape=Country), size=2) +
    scale_y_discrete(limits=rev) +
    xlab('Score on full data relative to 10k sample') +
    ylab('Dataset (country and sex)') +
    geom_vline(xintercept=0) +
    theme_bw() +
    theme(legend.position='bottom')
```























































