---
title: "varimp_effect_seed"
author: "Marieke Vinkenoog"
date: '2022-05-09'
output: html_document
---

# Introduction

In this script, we have 10 runs of the container on the same 10k donor subsample,
ran with different seeds. We compare the performance of these different runs - 
if there is no effect of the seed on the model fitting, the runs should all 
have the exact same performance.

By loading the fitted models and validation sets and
recalculating SHAP values with the same seed, we can see if there is an effect
of the seed anywhere else than in the 10k subsample selection (see performance_seeds.Rmd)
and in the selection of donors for calculating the SHAP values (see shap_effect_seed.Rmd).

# Libraries
```{r}
library(ggstance)
library(ggplot2)
library(dplyr)


library(Rcpp)
library(cutpointr)
library(farver)
library(PRROC)
library(future)
library(DALEX)
library(shapr)
library(themis)
library(caret)
library(kernlab)
library(tictoc)
library(forcats)
library(grid)
suppressPackageStartupMessages(library(parallel))
suppressPackageStartupMessages(library(doParallel))
suppressPackageStartupMessages(library(shapr))
suppressPackageStartupMessages(library(fastshap))
suppressPackageStartupMessages(library(ggforce))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(ranger))
```


# Functions to load data

```{r}
load_files <- function(folder, target_name) {
  info <- str_split(folder, '-', simplify=TRUE)       # seed, country, size, mode, version
  file <- paste0(path, folder, target_name)
  data <- read.csv(file) %>%
    mutate(seed = info[1],
           country = info[2],
           size = info[3],
           mode = info[4])
  return(data)
}

load_model_validate <- function(folder, path, modelname, sex) {
  info <- str_split(folder, '-', simplify=TRUE)
  fileloc <- paste0(path, folder, '/tmp_rds/')

  model <- readRDS(paste0(fileloc, paste0(modelname, '-fit-', sex, '.rds')))
  validate <- readRDS(paste0(fileloc, paste0(modelname, '-validate-', sex, '.rds')))
  
  both <- list(name=paste(folder, modelname, sex, sep='-'), model=model, validate=validate)
  
  return(both)
}
```

# Functions to calculate SHAP values
```{r}
# Caret's predict method only allows type to be "prob" or "raw.
# We will dig out the ksvm fit object from Caret's fit object and
# call its predict method directly.
my_predict <- function(fit, newdata, type="response") {
  stopifnot("train" %in% class(fit) && "ksvm" %in% class(fit$finalModel))
  newdata <- predict(fit$preProcess, newdata)  # Preprocess the data the same way Caret does
  newdata <- newdata %>% select(-any_of(c("sex", "Hb", "Hb_deferral"))) 
  res <- predict(fit$finalModel, newdata=newdata, type=type) # Call underlying object's predict method
  return(res)
}

compute_shap_values_fastshap <- function(result, seed=123, n=50, nsim=50) {
  set.seed(seed)
  name <- result$name
  model_seed <- str_split(name, '-', simplify=TRUE)[1]
  model <- result$model
  validate <- result$validate
  
  n <- min(n, nrow(validate))
  
  validate2 <- validate  %>% slice_sample(n=n) %>% select(-any_of(c("Hb_deferral", "Hb")))

  if ("stanfit" %in% class(model)) {
    t <- as_tibble(rownames_to_column(as.data.frame(rstan::get_posterior_mean(model))))
    beta <- t %>% filter(str_detect(rowname, r"(^beta\[\d+\])")) %>% pull(`mean-all chains`) #select(rowname, mean=`mean-all chains`)
  }
  
  pfun_lmm <- function(object, newdata) {
    #message(colnames(newdata))
    #message(head(newdata$donb))
    #message(sprintf("In function pfun_lmm: rows=%i cols=%i", nrow(newdata), ncol(newdata)))
    result <- as.vector(beta %*% t(as.matrix(newdata %>% select(-donb))) + newdata %>% pull(donb))
    return(result)
  }
  
  pfun_randomForest <- function(object, newdata) {
    predict(object, newdata = newdata, type="prob")[,2]
  }
  
  pfun_ranger <- function(object, newdata) {
    predict(object, data = newdata, type="response")$predictions[,"Deferred"]
  }

  # Not used currently as this doesn't perform preprocessing on newdata
  pfun_ksvm <- function(object, newdata) {
    predict(object, newdata = newdata, type="probabilities")[,2]
  }

  # This uses decision values instead of probabilities
  pfun_ksvm_decision <- function(object, newdata) {
    my_predict(object, newdata = newdata, type="decision")[,1]
  }
  
  # This is Caret's wrapper model
  pfun_train <- function(object, newdata) {
    predict(object, newdata = newdata, type="prob")[,2]
  }
  
  
  if ("randomForest" %in% class(model)) {
    pfun <- pfun_randomForest
  } else if ("ksvm" %in% class(model)) {
    pfun <- pfun_ksvm
  } else if ("ranger" %in% class(model)) {
    pfun <- pfun_ranger
  } else if ("train" %in% class(model)) {
    if ("ksvm" %in% class(model$finalModel)) {
      pfun <- pfun_ksvm_decision
    } else pfun <- pfun_train
  } else if ("stanfit" %in% class(model)) {
    pfun <- pfun_lmm
  }
  
  #print(pfun)
  
  result_code <- tryCatch(
    error = function(cnd) {
      msg <- paste("\nComputation of shap values failed:", cnd$message, 
                   sep="\n")
      warning(msg)
      NULL
    },
    {
      rlang::with_options(lifecycle_verbosity = "quiet", {  # Prevent the deprecation message caused by the explain function
        if ("lm" %in% class(model)) {
          shap <- fastshap::explain(model,   # This is for the baseline logistic regression
                                    feature_names = "previous_Hb",
                                    newdata = as.data.frame(validate2 %>% select(previous_Hb)),
                                    exact = TRUE)
        } else if ("stanfit" %in% class(model)) {
          shap <- fastshap::explain(model, 
                                    X = as.data.frame(validate2), 
                                    #newdata = as.data.frame(validate), 
                                    feature_names = setdiff(colnames(validate2), "donb"),
                                    pred_wrapper = pfun, 
                                    nsim = nsim)
        } else {
          shap <- fastshap::explain(model, 
                                    X = as.data.frame(validate2),
                                    pred_wrapper = pfun, 
                                    nsim = nsim)
        }
      })
      shap <- as_tibble(shap)  # This drops the class "explain"
    }
    
  )
  if (is.null(result_code)) {
    return(NULL)
  }
  
  if (any(shap %>% mutate(c=if_any(everything(), is.na)) %>% pull(c)) || nrow(shap) == 2*n) {
    warning("Predict function failed in compute_shap_values_fastshap. You could try rerunning with a different seed")
    return(NULL)
  }
  n <- nrow(validate2)
  attributions <- shap %>% mutate(id=1:n)
  attributions <- pivot_longer(attributions, cols=!id) %>%
    select(Variable=name, id, attribution=value)
  features <- validate2 %>% mutate(id=1:n)
  if ("sex" %in% colnames(validate2)) {
    features <- features %>% mutate(sex = sex=="female")
  }
  features <- features %>% pivot_longer(cols=!id) %>%
    select(Variable=name, id, value=value)
  res <- inner_join(attributions, features, by=c("Variable", "id")) %>%
    group_by(Variable) %>%
    mutate(value=scale(value)[,1]) %>%
    ungroup()
  
  res <- res %>% select(-id) %>% slice_sample(prop = 1.0)  # permute the rows
  res <- res %>% mutate(model_seed = model_seed)
  return(res)
}
```



# Functions to calculate and plot mean absolute attribution
```{r}
calc_mean_abs_attr <- function(df) {
  df <- df %>%
    filter(Variable != 'sex') %>%
    mutate(abs_attr = abs(attribution)) %>%
    group_by(model_seed, Variable) %>%
    summarise(mean_abs_attr = mean(abs_attr)) %>%
    ungroup() %>%
    mutate(Variable = as.factor(Variable),
           Variable = fct_relevel(Variable, 'previous_Hb', 'recent_deferrals',
                                  'Hb_first', 'days_to_previous_fb', 'recent_donations',
                                  'hour', 'previous_Hb_def', 'consecutive_deferrals',
                                  'warm_season', 'age'))
  return(list(df))
}

plot_maa_seeds <- function(df, title) {
  ggplot() +
    geom_point(data=df, aes(x=mean_abs_attr, y=Variable, color=model_seed)) +
    scale_colour_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                                 '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                                 '#cab2d6','#6a3d9a')) +
    scale_y_discrete(limits=rev) +
    ggtitle(paste('Mean absolute attribution in different model seeds,', title)) +
    theme_bw() + 
    theme(legend.position='none')
}
```



# Loading data
The summary.csv files contain the performance measures.

```{r}
path <- '../../results/multiple dutch 10k runs same sample/'

folders <- list.dirs(path, full.names=FALSE, recursive=FALSE)
folders <- folders[folders != '']

datalist <- lapply(folders, load_files, target_name='/summary.csv')
data <- bind_rows(datalist) %>%
  filter(Model %in% c('bl', 'rf', 'svm')) %>%
  rename(sex = Sex) %>%
  mutate(run = paste(size, seed, mode, sep='-'),
         run = str_replace(run, '10k-', ''))
```

# Comparing performance between different seeds

```{r}
data %>% 
  ggplot(aes(x=AUPR.value, xmin=AUPR.low, xmax=AUPR.high, y=run)) +
    facet_grid(rows=vars(Pretty), cols=vars(sex)) +
    ggstance::geom_pointrangeh(aes(colour=seed, shape=mode)) +
    scale_colour_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                               '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                               '#cab2d6','#6a3d9a')) +
    ylab('Dataset (seed and mode)') +
    theme_bw() +
    theme(legend.position='bottom')

ggsave('../effect_seed_modelfitting_plots/AUPR.png')
```
In the baseline model, there is no difference in performance between the 
different seeds. The only difference is due to the mode: initial has a higher
performance than final.

In the random forest (RF) and support vector machine (SVM) models, there are 
pronounced differences. Variation in performance is larger in men than in 
women, and is larger in the initial runs than in the final runs. The last
observation might be explained by the larger sample size in the final run,
which increases stability of estimates.

# Calculating variable importances in different runs -- mode initial

Because the data is exactly the same in every run (at least within the same mode),
the seed must be used during the training of the model. We calculate the
SHAP values of the different models to analyse the differences. We do this
outside the container rather than loading the shap-value.csv files, because
the SHAP values are calculated on a 1000 donor subsample, which is dependent on
the seed, and therefore a different subsample for every run. Here, we use the
same seed for all runs. 

```{r}
folders_initial <- grep('initial', folders, value=TRUE)

svm_female <- lapply(folders_initial, load_model_validate, path=path, model='svm', sex='female')
res_svm_female <- bind_rows(lapply(svm_female, compute_shap_values_fastshap, seed=123))
maa_svm_female <- calc_mean_abs_attr(res_svm_female)[[1]]

svm_male <- lapply(folders_initial, load_model_validate, path=path, model='svm', sex='male')
res_svm_male <- bind_rows(lapply(svm_male, compute_shap_values_fastshap, seed=123))
maa_svm_male <- calc_mean_abs_attr(res_svm_male)[[1]]

rf_female <- lapply(folders_initial, load_model_validate, path=path, model='rf', sex='female')
res_rf_female <- bind_rows(lapply(rf_female, compute_shap_values_fastshap, seed=123))
maa_rf_female <- calc_mean_abs_attr(res_rf_female)[[1]]

rf_male <- lapply(folders_initial, load_model_validate, path=path, model='rf', sex='male')
res_rf_male <- bind_rows(lapply(rf_male, compute_shap_values_fastshap, seed=123))
maa_rf_male <- calc_mean_abs_attr(res_rf_male)[[1]]

saveRDS(maa_svm_female, '../../results/multiple dutch 10k runs same sample/maa_svm_female_initial.RDS')
saveRDS(maa_rf_male, '../../results/multiple dutch 10k runs same sample/maa_svm_male_initial.RDS')
saveRDS(maa_svm_female, '../../results/multiple dutch 10k runs same sample/maa_rf_female_initial.RDS')
saveRDS(maa_rf_male, '../../results/multiple dutch 10k runs same sample/maa_rf_male_initial.RDS')
```

# Plots -- mode initial

```{r}
maa_svm_female <- readRDS('../../results/multiple dutch 10k runs same sample/maa_svm_female_initial.RDS')
maa_svm_male <- readRDS('../../results/multiple dutch 10k runs same sample/maa_svm_male_initial.RDS')
maa_rf_female <- readRDS('../../results/multiple dutch 10k runs same sample/maa_rf_female_initial.RDS')
maa_rf_male <- readRDS('../../results/multiple dutch 10k runs same sample/maa_rf_male_initial.RDS')

plot_maa_seeds(maa_svm_female, 'SVM female, initial')
ggsave('../effect_seed_modelfitting_plots/svm_female_initial.png')
plot_maa_seeds(maa_svm_male, 'SVM male, initial')
ggsave('../effect_seed_modelfitting_plots/svm_male_initial.png')
plot_maa_seeds(maa_rf_female, 'RF female, initial')
ggsave('../effect_seed_modelfitting_plots/rf_female_initial.png')
plot_maa_seeds(maa_rf_male, 'RF male, initial')
ggsave('../effect_seed_modelfitting_plots/rf_male_initial.png')
```

# Same for mode final

```{r}
folders_final <- grep('final', folders, value=TRUE)

svm_female <- lapply(folders_final, load_model_validate, path=path, model='svm', sex='female')
res_svm_female <- bind_rows(lapply(svm_female, compute_shap_values_fastshap, seed=123))
maa_svm_female <- calc_mean_abs_attr(res_svm_female)[[1]]

svm_male <- lapply(folders_final, load_model_validate, path=path, model='svm', sex='male')
res_svm_male <- bind_rows(lapply(svm_male, compute_shap_values_fastshap, seed=123))
maa_svm_male <- calc_mean_abs_attr(res_svm_male)[[1]]

rf_female <- lapply(folders_final, load_model_validate, path=path, model='rf', sex='female')
res_rf_female <- bind_rows(lapply(rf_female, compute_shap_values_fastshap, seed=123))
maa_rf_female <- calc_mean_abs_attr(res_rf_female)[[1]]

rf_male <- lapply(folders_final, load_model_validate, path=path, model='rf', sex='male')
res_rf_male <- bind_rows(lapply(rf_male, compute_shap_values_fastshap, seed=123))
maa_rf_male <- calc_mean_abs_attr(res_rf_male)[[1]]

saveRDS(maa_svm_female, '../../results/multiple dutch 10k runs same sample/maa_svm_female_final.RDS')
saveRDS(maa_rf_male, '../../results/multiple dutch 10k runs same sample/maa_svm_male_final.RDS')
saveRDS(maa_svm_female, '../../results/multiple dutch 10k runs same sample/maa_rf_female_final.RDS')
saveRDS(maa_rf_male, '../../results/multiple dutch 10k runs same sample/maa_rf_male_final.RDS')
```

# Plots -- mode final

```{r}
maa_svm_female <- readRDS('../../results/multiple dutch 10k runs same sample/maa_svm_female_final.RDS')
maa_svm_male <- readRDS('../../results/multiple dutch 10k runs same sample/maa_svm_male_final.RDS')
maa_rf_female <- readRDS('../../results/multiple dutch 10k runs same sample/maa_rf_female_final.RDS')
maa_rf_male <- readRDS('../../results/multiple dutch 10k runs same sample/maa_rf_male_final.RDS')

plot_maa_seeds(maa_svm_female, 'SVM female, final')
ggsave('../effect_seed_modelfitting_plots/svm_female_final.png')
plot_maa_seeds(maa_svm_male, 'SVM male, final')
ggsave('../effect_seed_modelfitting_plots/svm_male_final.png')
plot_maa_seeds(maa_rf_female, 'RF female, final')
ggsave('../effect_seed_modelfitting_plots/rf_female_final.png')
plot_maa_seeds(maa_rf_male, 'RF male, final')
ggsave('../effect_seed_modelfitting_plots/rf_male_final.png')
```


























