---
title: "Comparing multiple subsamples"
author: "Marieke Vinkenoog"
date: "4/8/2022"
output: html_document
---

# Introduction

This script looks at how different seeds affect performance (AUPR) of models
in the Hb-predictor container. Runs with ten different seeds were performed
on the Dutch data, with 10k subsamples. The seed affects the selection of the
subsample, which is the most obvious cause for variation in performance.
The seed might also be used by packages when the models are trained, and thus
lead to different models even if the same data is used (see model_effect_seed.Rmd).
Last, the seed is used by the SHAP value calculator to take a 100 donor subsample
to calculate SHAP values on (see shap_effect_seed.Rmd).

Here, we look at the variation in performance in runs with 10 different seeds.


# Libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stringr)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(ggpubr)
```

# Functions for loading data
```{r}
# For summary files containing model performances
load_files <- function(folder, target_name) {
  info <- str_split(folder, '-', simplify=TRUE)       # seed, country, size, mode, version
  file <- paste0(path, folder, target_name)
  data <- read.csv(file) %>%
    mutate(seed = info[1],
           country = info[2],
           size = info[3],
           mode = info[4])
  return(data)
}

# Called by load_dataframes below
load_dataframes_inner <- function(target_name, folder, info) {
  info2 = str_split(target_name, '-', simplify=TRUE)
  file <- paste0(path, folder, '/tmp_rds', target_name)
  data <- readRDS(file) %>%
    mutate(seed = info[1],
           country = info[2],
           size = info[3],
           mode = info[4],
           split = info2[2],
           sex = info2[3])
  return(data)
}

# For fitted model and validation files (not usually returned by container!)
load_dataframes <- function(folder, target_names) {
  info <- str_split(folder, '-', simplify=TRUE)       # seed, country, size, mode, version
  datalist <- lapply(target_names, load_dataframes_inner, folder=folder, info=info)
  data <- bind_rows(datalist)
  return(data)
}
```

# Load summary data

```{r}
path <- '../../results/multiple dutch 10k runs v2/'

folders <- list.dirs(path, full.names=FALSE, recursive=FALSE)
folders <- folders[folders != '']

# Data on model performances
datalist <- lapply(folders, load_files, target_name='/summary.csv')
data <- bind_rows(datalist) %>%
  filter(Model %in% c('bl', 'rf', 'svm')) %>%
  rename(ID = Id,
         sex = Sex) %>%
  mutate(run = paste(size, seed, mode, sep='-'),
         run = str_replace(run, '10k-', ''))

# Add corresponding AUPR on full data as a separate variable
for (row in 1:nrow(data)) {
  if (data[row, 'size'] == 'full') {
    data[row, 'AUPR.full'] = data[row, 'AUPR.value']
  } else {
    cur_mode = data[row, 'mode']
    cur_ID = data[row, 'ID']
    data[row, 'AUPR.full'] = data[(data$mode == cur_mode) & 
                                  (data$ID == cur_ID) &
                                  (data$size == 'full'), 'AUPR.value']
  }
}
```

# Plot AUPR differences between runs

```{r}
data %>% 
  ggplot(aes(x=AUPR.value, xmin=AUPR.low, xmax=AUPR.high, y=run)) +
    facet_grid(rows=vars(Pretty), cols=vars(sex)) +
    ggstance::geom_pointrangeh(aes(colour=seed, shape=mode)) +
    scale_colour_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                               '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                               '#cab2d6','#6a3d9a')) +
    geom_vline(aes(xintercept=AUPR.full, linetype=mode)) +
    ylab('Dataset (seed and mode)') +
    theme_bw() +
    theme(legend.position='bottom')
ggsave('../effect_10ksample_performance_plots/AUPR.png')
```

# Plot difference between AUPR in both modes between runs

```{r}
data_paired <- data %>%
  select(ID, Model, Pretty, sex, AUPR.value, seed, mode, size) %>%
  pivot_wider(names_from=mode, values_from=AUPR.value) %>%
  mutate(AUPR.diff = final - initial,
         Run = paste(size, seed, sep='-'),
         Run = str_replace(Run, '10k-', ''))

data_paired %>% 
  ggplot(aes(x=AUPR.diff, y=Run)) +
    facet_grid(rows=vars(Pretty), cols=vars(sex)) +
    geom_point(aes(colour=seed)) +
    scale_colour_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                               '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                               '#cab2d6','#6a3d9a')) +
    geom_vline(xintercept=0) +
    xlab('AUPR difference (final minus initial)') +
    ylab('Seed') +
    theme_bw() +
    theme(legend.position='bottom')

ggsave('../effect_10ksample_performance_plots/AUPR_mode_diff.png')
```

# Load fitted models and validation files 
```{r}
dflist <- lapply(folders, load_dataframes, target_names=c('/bl-train-male.rds', 
                                                          '/bl-train-female.rds',
                                                          '/bl-validate-male.rds',
                                                          '/bl-validate-female.rds'))
dfs <- bind_rows(dflist) %>%
  select(seed, size, mode, split, sex, age, days_to_previous_fb, previous_Hb_def,
         warm_season, consecutive_deferrals, recent_donations, recent_deferrals, 
         hour, previous_Hb, Hb_first, Hb_deferral) %>%
  mutate(run = paste(size, seed, mode, sep='-'),
         run = str_replace(run, '10k-', '')) %>%
  separate(sex, c('sex', 'removeme')) %>%
  select(-removeme)
```


# Comparing marginal distributions of predictor variables 

Men
```{r}
dfs %>%
  filter(sex == 'male') %>%
  ggplot() +
  facet_grid(split~mode) +
  geom_density(aes(x=previous_Hb, color=run)) +
  ggtitle('Distribution of previous_Hb by run (male donors)')
ggsave('../effect_10ksample_performance_plots/distributions_prevHb_male_nozoom.png', width=10, height=10)

dfs %>%
  filter(sex == 'female') %>%
  ggplot() +
  facet_grid(split~mode) +
  geom_density(aes(x=previous_Hb, color=run)) +
  ggtitle('Distribution of previous_Hb by run (female donors)')
ggsave('../effect_10ksample_performance_plots/distributions_prevHb_female_nozoom.png', width=10, height=10)

dfs %>%
  filter(sex == 'male') %>%
  ggplot() +
  facet_grid(split~mode) +
  geom_density(aes(x=previous_Hb, color=run)) +
  xlim(110,190) +
  ggtitle('Distribution of previous_Hb by run (male donors)')
ggsave('../effect_10ksample_performance_plots/distributions_prevHb_male_zoom.png', width=10, height=10)

dfs %>%
  filter(sex == 'female') %>%
  ggplot() +
  facet_grid(split~mode) +
  geom_density(aes(x=previous_Hb, color=run)) +
  xlim(100,180) +
  ggtitle('Distribution of previous_Hb by run (female donors)')
ggsave('../effect_10ksample_performance_plots/distributions_prevHb_female_zoom.png', width=10, height=10)
```


# Comparing variable importances (based on SHAP values) across runs

```{r}
datalist <- lapply(folders, load_files, target_name='/shap-value.csv')
varimps <- bind_rows(datalist) %>%
  rename(ID = Id) %>%
  mutate(run = paste(size, seed, mode, sep='-')) %>%
  separate(ID, c('model', 'sex')) %>%
  group_by(model, Variable, sex, run) %>%
  summarise(maa = mean(abs(attribution))) %>%
  ungroup() %>%
  separate(run, c('size', 'seed', 'mode'), remove=FALSE) %>%
  mutate(run = str_replace(run, '10k-', '')) 
```

# Plots RF
```{r}
varimps %>%
  filter(sex == 'male',
         model == 'rf',
         Variable != 'sex') %>%
  ggplot() +
  facet_wrap(vars(Variable)) +
  geom_col(aes(x=run, y=maa, fill=seed), color='black') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle('Mean absolute attribution by run for Random Forest models, male donors') +
  ylab('Mean absolute attribution') +
  scale_fill_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                               '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                               '#cab2d6','#6a3d9a'))
ggsave('../effect_10ksample_performance_plots/MAA_rf_male.png', width=10, height=10)

varimps %>%
  filter(sex == 'female',
         model == 'rf',
         Variable != 'sex') %>%
  ggplot() +
  facet_wrap(vars(Variable)) +
  geom_col(aes(x=run, y=maa, fill=seed), color='black') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle('Mean absolute attribution by run for Random Forest models, female donors') +
  ylab('Mean absolute attribution') +
  scale_fill_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                               '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                               '#cab2d6','#6a3d9a'))
ggsave('../effect_10ksample_performance_plots/MAA_rf_female.png', width=10, height=10)
```
# Plots SVM
```{r}
varimps %>%
  filter(sex == 'male',
         model == 'svm',
         Variable != 'sex') %>%
  ggplot() +
  facet_wrap(vars(Variable)) +
  geom_col(aes(x=run, y=maa, fill=seed), color='black') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle('Mean absolute attribution by run for SVM models, male donors') +
  ylab('Mean absolute attribution') +
  scale_fill_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                               '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                               '#cab2d6','#6a3d9a'))
ggsave('../effect_10ksample_performance_plots/MAA_svm_male.png', width=10, height=10)

varimps %>%
  filter(sex == 'female',
         model == 'svm',
         Variable != 'sex') %>%
  ggplot() +
  facet_wrap(vars(Variable)) +
  geom_col(aes(x=run, y=maa, fill=seed), color='black') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle('Mean absolute attribution by run for SVM models, female donors') +
  ylab('Mean absolute attribution') +
  scale_fill_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                               '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                               '#cab2d6','#6a3d9a'))
ggsave('../effect_10ksample_performance_plots/MAA_svm_female.png', width=10, height=10)
```

# Plots previous_Hb all models
```{r}
varimps %>%
  filter(Variable == 'previous_Hb') %>%
  ggplot() +
  facet_grid(rows=vars(sex), cols=vars(model)) +
  geom_col(aes(x=run, y=maa, fill=seed), color='black') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle('Mean absolute attribution of previous_Hb by run for all models, both donors') +
  ylab('Mean absolute attribution') +
  scale_fill_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                               '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                               '#cab2d6','#6a3d9a'))
ggsave('../effect_10ksample_performance_plots/MAA_all_prevHb.png', width=10, height=10)


```


# Scatterplot of MAA in RF versus SVM

```{r}
varimps_wide <- varimps %>%
  pivot_wider(names_from = model, values_from = maa)

varimps_wide %>%
  filter(sex == 'male',
         Variable != 'sex',
         mode == 'final') %>%
  ggplot() +
  facet_wrap(vars(Variable)) +
  geom_point(aes(x=rf, y=svm, color=seed)) + 
  scale_color_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                               '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                               '#cab2d6','#6a3d9a')) +
  ggtitle('MAA of random forest vs support vector machine, male donors') +
  theme_bw()
ggsave('../effect_10ksample_performance_plots/MAA_rf_svm_male.png', width=10, height=10)

varimps_wide %>%
  filter(sex == 'female',
         Variable != 'sex',
         mode == 'final') %>%
  ggplot() +
  facet_wrap(vars(Variable)) +
  geom_point(aes(x=rf, y=svm, color=seed)) + 
  scale_color_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                               '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                               '#cab2d6','#6a3d9a')) +
  ggtitle('MAA of random forest vs support vector machine, female donors') +
  theme_bw()
ggsave('../effect_10ksample_performance_plots/MAA_rf_svm_female.png', width=10, height=10)
```

# Scaled by total absolute attribution

```{r}
varimps_wide <- varimps_wide %>%
  group_by(sex, run) %>%
  mutate(rf_scaled = rf / sum(rf),
         svm_scaled = svm / sum(svm))

varimps_wide %>%
  filter(sex == 'male',
         Variable != 'sex',
         mode == 'final') %>%
  ggplot() +
  facet_wrap(vars(Variable), scales='free') +
  geom_point(aes(x=rf_scaled, y=svm_scaled, color=seed)) + 
  scale_color_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                               '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                               '#cab2d6','#6a3d9a')) +
  ggtitle('MAA of random forest vs support vector machine, male donors') +
  theme_bw()
ggsave('../multiple 10k runs/MAA_scaled_rf_svm_male.png', width=10, height=10)

varimps_wide %>%
  filter(sex == 'female',
         Variable != 'sex',
         mode == 'final') %>%
  ggplot() +
  facet_wrap(vars(Variable), scales='free') +
  geom_point(aes(x=rf_scaled, y=svm_scaled, color=seed)) + 
  scale_color_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                               '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                               '#cab2d6','#6a3d9a')) +
  ggtitle('MAA of random forest vs support vector machine, female donors') +
  theme_bw()
ggsave('../multiple 10k runs/MAA_scaled_rf_svm_female.png', width=10, height=10)
```


# Investigate number of deferrals in sample vs AUPR

```{r}
datalist <- lapply(folders, load_files, target_name='/sizes.csv')
data_sizes <- bind_rows(datalist) %>%
  rename(ID = Id) %>%
  mutate(run = paste(size, seed, mode, sep='-'),
         run = str_replace(run, '10k-', '')) %>%
  separate(ID, c('model', 'sex')) %>%
  filter(model == 'bl') %>%
  group_by(sex, label, run) %>%
  mutate(ndeferrals = sum(`Deferred.last.donations`)) %>%
  ungroup() %>%
  select(c(sex, label, seed, size, mode, run, ndeferrals)) %>%
  distinct() %>%
  pivot_wider(names_from=label, values_from=ndeferrals) %>%
  mutate(ndeferrals_train = case_when(mode == 'initial' ~ train,
                                      mode == 'final' ~ train+validate),
         ndeferrals_test = case_when(mode == 'initial' ~ validate,
                                     mode == 'final' ~ test),
         defrate_train = case_when(mode == 'initial' ~ ndeferrals_train/6400,
                                   mode == 'final' ~ ndeferrals_train/8000),
         defrate_test = case_when(mode == 'initial' ~ ndeferrals_test/1600,
                                  mode == 'final' ~ ndeferrals_test/2000))

data <- data %>%
  merge(data_sizes, by=c('sex', 'run', 'seed', 'size', 'mode')) %>%
  select(-starts_with(c('MAE', 'RMSE', 'AUROC', 'F1')))
```

# Plot proportion of deferrals vs AUPR

```{r}
data %>%
  filter(size == '10k') %>%
  ggplot() +
  facet_grid(sex ~ Model) +
  geom_point(aes(x=defrate_train, y=AUPR.value, color=seed, shape=mode), size=2) +
  scale_colour_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                               '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                               '#cab2d6','#6a3d9a')) +
  theme_bw() +
  ggtitle('Proportion of deferrals in train data versus AUPR')
ggsave('../effect_10ksample_performance_plots/defrate_train_AUPR.png', width=10, height=10)

data %>%
  filter(size == '10k') %>%
  ggplot() +
  facet_grid(sex ~ Model) +
  geom_point(aes(x=defrate_test, y=AUPR.value, color=seed, shape=mode), size=2) +
  scale_colour_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                               '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                               '#cab2d6','#6a3d9a')) +
  theme_bw() +
  ggtitle('Proportion of deferrals in test data versus AUPR')
ggsave('../effect_10ksample_performance_plots/defrate_test_AUPR.png', width=7, height=4)
```














