---
title: "Comparing multiple subsamples"
author: "Marieke Vinkenoog"
date: "4/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stringr)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(ggpubr)
```

## Loading data


```{r}
#path <- ''   # Point this to where you saved the folders from all-runs.zip
path <- 'C:/Users/mvink/OneDrive/PhD/Research/202105 SanguinStats Hb deferral/results/multiple dutch 10k runs v2/'
setwd(path)

folders <- list.dirs(path, full.names=FALSE, recursive=FALSE)
folders <- folders[folders != '']

load_files <- function(folder, target_name) {
  info <- str_split(folder, '-', simplify=TRUE)       # seed, country, size, mode, version
  file <- paste0(path, folder, target_name)
  data <- read.csv(file) %>%
    mutate(seed = info[1],
           country = info[2],
           size = info[3],
           mode = info[4])
  return(data)
}

load_dataframes_inner <- function(target_name, folder, info) {
  info2 = str_split(target_name, '-', simplify=TRUE)
  file <- paste0(path, folder, '/tmp_rds', target_name)
  data <- readRDS(file) %>%
    mutate(seed = info[1],
           country = info[2],
           size = info[3],
           mode = info[4],
           split = info2[2],
           sex = info2[3])
  return(data)
}

load_dataframes <- function(folder, target_names) {
  info <- str_split(folder, '-', simplify=TRUE)       # seed, country, size, mode, version
  datalist <- lapply(target_names, load_dataframes_inner, folder=folder, info=info)
  data <- bind_rows(datalist)
  return(data)
}

datalist <- lapply(folders, load_files, target_name='/summary.csv')
data <- bind_rows(datalist) %>%
  filter(Model %in% c('bl', 'rf', 'svm')) %>%
  rename(ID = `誰..Id`,
         sex = Sex) %>%
  mutate(run = paste(size, seed, mode, sep='-'),
         run = str_replace(run, '10k-', ''))

for (row in 1:nrow(data)) {
  if (data[row, 'size'] == 'full') {
    data[row, 'AUPR.full'] = data[row, 'AUPR.value']
  } else {
    cur_mode = data[row, 'mode']
    cur_ID = data[row, 'ID']
    data[row, 'AUPR.full'] = data[(data$mode == cur_mode) & 
                                  (data$ID == cur_ID) &
                                  (data$size == 'full'), 'AUPR.value']
  }
}

dflist <- lapply(folders, load_dataframes, target_names=c('/bl-train-male.rds', 
                                                          '/bl-train-female.rds',
                                                          '/bl-validate-male.rds',
                                                          '/bl-validate-female.rds'))
dfs <- bind_rows(dflist) %>%
  select(seed, size, mode, split, sex, age, days_to_previous_fb, previous_Hb_def,
         warm_season, consecutive_deferrals, recent_donations, recent_deferrals, 
         hour, previous_Hb, Hb_first, Hb_deferral) %>%
  mutate(run = paste(size, seed, mode, sep='-'),
         run = str_replace(run, '10k-', '')) %>%
  separate(sex, c('sex', 'removeme')) %>%
  select(-removeme)
```

```{r}
data %>% 
  ggplot(aes(x=AUPR.value, xmin=AUPR.low, xmax=AUPR.high, y=Run)) +
    facet_grid(rows=vars(Pretty), cols=vars(Sex)) +
    ggstance::geom_pointrangeh(aes(colour=seed, shape=mode)) +
    scale_colour_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                               '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                               '#cab2d6','#6a3d9a')) +
    geom_vline(aes(xintercept=AUPR.full, linetype=mode)) +
    ylab('Dataset (seed and mode)') +
    theme_bw() +
    theme(legend.position='bottom')
```

```{r}
data_paired <- data %>%
  select(ID, Model, Pretty, sex, AUPR.value, seed, mode, size) %>%
  pivot_wider(names_from=mode, values_from=AUPR.value) %>%
  mutate(AUPR.diff = final - initial,
         Run = paste(size, seed, sep='-'),
         Run = str_replace(Run, '10k-', ''))

data_paired %>% 
  ggplot(aes(x=AUPR.diff, y=Run)) +
    facet_grid(rows=vars(Pretty), cols=vars(sex)) +
    geom_point(aes(colour=seed)) +
    scale_colour_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                               '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                               '#cab2d6','#6a3d9a')) +
    geom_vline(xintercept=0) +
    xlab('AUPR difference (final minus initial)') +
    ylab('Seed') +
    theme_bw() +
    theme(legend.position='bottom')
```

# Load histogram files
(old version, replaced by histograms from raw data)

```{r}
datalist <- lapply(folders, load_files, target_name='/histogram.csv')
distrs <- bind_rows(datalist) %>%
  rename(variable = `誰..name`) %>%
  separate(id, c(NA, 'sex')) %>%
  separate(value, c(NA, 'val.lower', 'val.upper', NA), sep='[\\(\\],]', remove=FALSE, convert=TRUE) %>%
  filter(sex != 'both',
         variable != 'Year') %>%
  distinct() %>%
  mutate(run = paste(size, seed, mode, sep='-'),
         run = str_replace(run, '10k-', ''),
         variable = as.factor(variable),
         sex = as.factor(sex),
         run = as.factor(run))

distrs %>% 
  filter(sex == 'female',
         !variable %in% c('first_event', 'nb_donat_outside', 'Warm season'),
         size == '10k',
         mode == 'initial') %>% 
  ggplot(aes(color=run)) +
    facet_wrap(~variable, scales='free') +
    geom_density(aes(x=val.lower, y=n), stat='identity') +
    geom_point(aes(x=val.lower, y=n)) +
    scale_colour_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                                 '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                                 '#cab2d6','#6a3d9a')) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
distrs %>% 
  filter(variable %in% c('First Hb', 'Hemoglobin', 'Previous Hb'),
         size == '10k',
         mode == 'initial') %>% 
  ggplot(aes(color=run)) +
    facet_grid(rows=vars(sex), cols=vars(variable), scales='free') +
    geom_density(aes(x=val.lower, y=n), stat='identity') +
    geom_point(aes(x=val.lower, y=n)) +
    scale_colour_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                                 '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                                 '#cab2d6','#6a3d9a')) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
distrs %>% 
  filter(variable == 'Previous Hb',
         size == '10k',
         mode == 'initial') %>% 
  ggplot(aes(color=run)) +
    facet_grid(rows=vars(sex), scales='free') +
    geom_density(aes(x=val.lower, y=n), stat='identity') +
    geom_point(aes(x=val.lower, y=n)) +
    scale_colour_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                                 '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                                 '#cab2d6','#6a3d9a')) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```


# Comparing marginal distributions of predictor variables using raw data

Men
```{r}
dfs %>%
  filter(sex == 'male') %>%
  ggplot() +
  facet_grid(split~mode) +
  geom_density(aes(x=previous_Hb, color=run)) +
  ggtitle('Distribution of previous_Hb by run (male donors)')
ggsave('../multiple 10k runs/distributions_prevHb_male_nozoom.png', width=10, height=10)
```
Women
```{r}
dfs %>%
  filter(sex == 'female') %>%
  ggplot() +
  facet_grid(split~mode) +
  geom_density(aes(x=previous_Hb, color=run)) +
  ggtitle('Distribution of previous_Hb by run (female donors)')
ggsave('../multiple 10k runs/distributions_prevHb_female_nozoom.png', width=10, height=10)
```

# Comparing variable importances (based on SHAP values) across runs

```{r}
datalist <- lapply(folders, load_files, target_name='/shap-value.csv')
varimps <- bind_rows(datalist) %>%
  rename(ID = `誰..Id`) %>%
  mutate(run = paste(size, seed, mode, sep='-')) %>%
  separate(ID, c('model', 'sex')) %>%
  group_by(model, Variable, sex, run) %>%
  summarise(maa = mean(abs(attribution))) %>%
  ungroup() %>%
  separate(run, c('size', 'seed', 'mode'), remove=FALSE) %>%
  mutate(run = str_replace(run, '10k-', '')) 
```

# Plots RF
```{r}
varimps %>%
  filter(sex == 'male',
         model == 'rf',
         Variable != 'sex') %>%
  ggplot() +
  facet_wrap(vars(Variable)) +
  geom_col(aes(x=run, y=maa, fill=seed), color='black') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle('Mean absolute attribution by run for Random Forest models, male donors') +
  ylab('Mean absolute attribution') +
  scale_fill_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                               '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                               '#cab2d6','#6a3d9a'))
ggsave('../multiple 10k runs/MAE_rf_male.png', width=10, height=10)

varimps %>%
  filter(sex == 'female',
         model == 'rf',
         Variable != 'sex') %>%
  ggplot() +
  facet_wrap(vars(Variable)) +
  geom_col(aes(x=run, y=maa, fill=seed), color='black') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle('Mean absolute attribution by run for Random Forest models, female donors') +
  ylab('Mean absolute attribution') +
  scale_fill_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                               '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                               '#cab2d6','#6a3d9a'))
ggsave('../multiple 10k runs/MAE_rf_female.png', width=10, height=10)
```
# Plots SVM
```{r}
varimps %>%
  filter(sex == 'male',
         model == 'svm',
         Variable != 'sex') %>%
  ggplot() +
  facet_wrap(vars(Variable)) +
  geom_col(aes(x=run, y=maa, fill=seed), color='black') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle('Mean absolute attribution by run for SVM models, male donors') +
  ylab('Mean absolute attribution') +
  scale_fill_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                               '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                               '#cab2d6','#6a3d9a'))
ggsave('../multiple 10k runs/MAE_svm_male.png', width=10, height=10)

varimps %>%
  filter(sex == 'female',
         model == 'svm',
         Variable != 'sex') %>%
  ggplot() +
  facet_wrap(vars(Variable)) +
  geom_col(aes(x=run, y=maa, fill=seed), color='black') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle('Mean absolute attribution by run for SVM models, female donors') +
  ylab('Mean absolute attribution') +
  scale_fill_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                               '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                               '#cab2d6','#6a3d9a'))
ggsave('../multiple 10k runs/MAE_svm_female.png', width=10, height=10)
```

# Plots previous_Hb all models
```{r}
varimps %>%
  filter(Variable == 'previous_Hb') %>%
  ggplot() +
  facet_grid(rows=vars(sex), cols=vars(model)) +
  geom_col(aes(x=run, y=maa, fill=seed), color='black') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle('Mean absolute attribution of previous_Hb by run for all models, both donors') +
  ylab('Mean absolute attribution') +
  scale_fill_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                               '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                               '#cab2d6','#6a3d9a'))
ggsave('../multiple 10k runs/MAE_all_prevHb.png', width=10, height=10)


```


# Scatterplot of MAA in RF versus SVM

```{r}
varimps_wide <- varimps %>%
  pivot_wider(names_from = model, values_from = maa)



varimps_wide %>%
  filter(sex == 'male',
         Variable != 'sex',
         mode == 'final') %>%
  ggplot() +
  facet_wrap(vars(Variable)) +
  geom_point(aes(x=rf, y=svm, color=seed)) + 
  scale_color_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                               '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                               '#cab2d6','#6a3d9a')) +
  ggtitle('MAA of random forest vs support vector machine, male donors') +
  theme_bw()
ggsave('../multiple 10k runs/MAA_rf_svm_male.png', width=10, height=10)

varimps_wide %>%
  filter(sex == 'female',
         Variable != 'sex',
         mode == 'final') %>%
  ggplot() +
  facet_wrap(vars(Variable)) +
  geom_point(aes(x=rf, y=svm, color=seed)) + 
  scale_color_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                               '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                               '#cab2d6','#6a3d9a')) +
  ggtitle('MAA of random forest vs support vector machine, female donors') +
  theme_bw()
ggsave('../multiple 10k runs/MAA_rf_svm_female.png', width=10, height=10)
```

# Scaled by total absolute attribution

```{r}
varimps_wide <- varimps_wide %>%
  group_by(sex, run) %>%
  mutate(rf_scaled = rf / sum(rf),
         svm_scaled = svm / sum(svm))

varimps_wide %>%
  filter(sex == 'male',
         Variable != 'sex',
         mode == 'final') %>%
  ggplot() +
  facet_wrap(vars(Variable), scales='free') +
  geom_point(aes(x=rf_scaled, y=svm_scaled, color=seed)) + 
  scale_color_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                               '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                               '#cab2d6','#6a3d9a')) +
  ggtitle('MAA of random forest vs support vector machine, male donors') +
  theme_bw()
ggsave('../multiple 10k runs/MAA_scaled_rf_svm_male.png', width=10, height=10)

varimps_wide %>%
  filter(sex == 'female',
         Variable != 'sex',
         mode == 'final') %>%
  ggplot() +
  facet_wrap(vars(Variable), scales='free') +
  geom_point(aes(x=rf_scaled, y=svm_scaled, color=seed)) + 
  scale_color_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                               '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                               '#cab2d6','#6a3d9a')) +
  ggtitle('MAA of random forest vs support vector machine, female donors') +
  theme_bw()
ggsave('../multiple 10k runs/MAA_scaled_rf_svm_female.png', width=10, height=10)
```


# Investigate number of deferrals in sample vs AUPR

```{r}
datalist <- lapply(folders, load_files, target_name='/sizes.csv')
data_sizes <- bind_rows(datalist) %>%
  rename(ID = `誰..Id`) %>%
  mutate(run = paste(size, seed, mode, sep='-'),
         run = str_replace(run, '10k-', '')) %>%
  separate(ID, c('model', 'sex')) %>%
  filter(model == 'bl') %>%
  group_by(sex, label, run) %>%
  mutate(ndeferrals = sum(`Deferred.last.donations`)) %>%
  ungroup() %>%
  select(c(sex, label, seed, size, mode, run, ndeferrals)) %>%
  distinct() %>%
  pivot_wider(names_from=label, values_from=ndeferrals) %>%
  mutate(ndeferrals_train = case_when(mode == 'initial' ~ train,
                                      mode == 'final' ~ train+validate),
         ndeferrals_test = case_when(mode == 'initial' ~ validate,
                                     mode == 'final' ~ test))

#data_no_sizes <- data 

data <- data_no_sizes %>%
  merge(data_sizes, by=c('sex', 'run', 'seed', 'size', 'mode')) %>%
  select(-starts_with(c('MAE', 'RMSE', 'AUROC', 'F1')))
```

# Plots ndeferrals vs AUPR

```{r}
data %>%
  filter(size == '10k') %>%
  ggplot() +
  facet_grid(sex ~ Model) +
  geom_point(aes(x=ndeferrals_train, y=AUPR.value, color=seed, shape=mode), size=2) +
  scale_colour_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                               '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                               '#cab2d6','#6a3d9a')) +
  theme_bw() +
  ggtitle('Number of deferrals in train data versus AUPR')
ggsave('../multiple 10k runs/ndeferrals_train_AUPR.png', width=10, height=10)

data %>%
  filter(size == '10k') %>%
  ggplot() +
  facet_grid(sex ~ Model) +
  geom_point(aes(x=ndeferrals_test, y=AUPR.value, color=seed, shape=mode), size=2) +
  scale_colour_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                               '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                               '#cab2d6','#6a3d9a')) +
  theme_bw() +
  ggtitle('Number of deferrals in test data versus AUPR')
ggsave('../multiple 10k runs/ndeferrals_test_AUPR.png', width=10, height=10)
```

# Plot ndeferral vs width of CI

```{r}
data %>%
  filter(size == '10k') %>%
  mutate(CIwidth = AUPR.high - AUPR.low) %>%
  ggplot() +
  facet_grid(sex ~ Model) +
  geom_point(aes(x=ndeferrals_train, y=CIwidth, color=seed, shape=mode), size=2) +
  scale_colour_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                               '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                               '#cab2d6','#6a3d9a')) +
  theme_bw() +
  ggtitle('Number of deferrals in train data versus CI width')
ggsave('../multiple 10k runs/ndeferrals_train_CIwidth.png', width=10, height=10)

data %>%
  filter(size == '10k') %>%
  mutate(CIwidth = AUPR.high - AUPR.low) %>%
  ggplot() +
  facet_grid(sex ~ Model) +
  geom_point(aes(x=ndeferrals_test, y=CIwidth, color=seed, shape=mode), size=2) +
  scale_colour_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                               '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                               '#cab2d6','#6a3d9a')) +
  theme_bw() +
  ggtitle('Number of deferrals in test data versus CI width')
ggsave('../multiple 10k runs/ndeferrals_test_CIwidth.png', width=10, height=10)
```

# AUPR versus CI width?

```{r}
data %>%
  filter(size == '10k') %>%
  mutate(CIwidth = AUPR.high - AUPR.low) %>%
  ggplot() +
  facet_grid(sex ~ Model) +
  geom_point(aes(x=AUPR.value, y=CIwidth, color=seed, shape=mode), size=2) +
  scale_colour_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                               '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                               '#cab2d6','#6a3d9a')) +
  theme_bw() +
  ggtitle('AUPR value versus CI width')
ggsave('../multiple 10k runs/AUPR_CIwidth.png', width=10, height=10)
```















