---
title: "Effect of seed on SHAP value computation"
author: "Marieke Vinkenoog"
date: "03/05/2022"
output: html_document
---

# Introduction

In this script we take one container run, and compute SHAP values multiple
times using different seeds to select the donors to compute them on.
We do this for n=1000 (used in the container), n=100 and n=2000 (the full
test set in the final run).

# Libraries

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(Rcpp)
library(cutpointr)
library(farver)
library(PRROC)
library(future)
library(DALEX)
library(shapr)
library(fastshap)
library(themis)
library(caret)
library(kernlab)
library(tictoc)
library(forcats)
library(grid)
suppressPackageStartupMessages(library(parallel))
suppressPackageStartupMessages(library(doParallel))
suppressPackageStartupMessages(library(shapr))
suppressPackageStartupMessages(library(fastshap))
suppressPackageStartupMessages(library(ggforce))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(ranger))
```


# Functions for loading data
```{r}
load_model_validate <- function(path, folder, modelname, sex) {
  info <- str_split(folder, '-', simplify=TRUE)
  fileloc <- paste0(path, folder, '/tmp_rds/')

  model <- readRDS(paste0(fileloc, paste0(modelname, '-fit-', sex, '.rds')))
  validate <- readRDS(paste0(fileloc, paste0(modelname, '-validate-', sex, '.rds')))
  
  both <- list(name=paste(folder, modelname, sex, sep='-'), model=model, validate=validate)
  
  return(both)
}
```


# Functions for calculating SHAP values (extracted from container)

```{r}
# Caret's predict method only allows type to be "prob" or "raw.
# We will dig out the ksvm fit object from Caret's fit object and
# call its predict method directly.
my_predict <- function(fit, newdata, type="response") {
  stopifnot("train" %in% class(fit) && "ksvm" %in% class(fit$finalModel))
  newdata <- predict(fit$preProcess, newdata)  # Preprocess the data the same way Caret does
  newdata <- newdata %>% select(-any_of(c("sex", "Hb", "Hb_deferral"))) 
  res <- predict(fit$finalModel, newdata=newdata, type=type) # Call underlying object's predict method
  return(res)
}

compute_shap_values_fastshap <- function(model, validate, n=1000, seed, nsim=100) {
  set.seed(seed)
  
  n <- min(n, nrow(validate))
  
  validate2 <- validate  %>% slice_sample(n=n) %>% select(-any_of(c("Hb_deferral", "Hb")))

  if ("stanfit" %in% class(model)) {
    t <- as_tibble(rownames_to_column(as.data.frame(rstan::get_posterior_mean(model))))
    beta <- t %>% filter(str_detect(rowname, r"(^beta\[\d+\])")) %>% pull(`mean-all chains`) #select(rowname, mean=`mean-all chains`)
  }
  
  pfun_lmm <- function(object, newdata) {
    #message(colnames(newdata))
    #message(head(newdata$donb))
    #message(sprintf("In function pfun_lmm: rows=%i cols=%i", nrow(newdata), ncol(newdata)))
    result <- as.vector(beta %*% t(as.matrix(newdata %>% select(-donb))) + newdata %>% pull(donb))
    return(result)
  }
  
  pfun_randomForest <- function(object, newdata) {
    predict(object, newdata = newdata, type="prob")[,2]
  }
  
  pfun_ranger <- function(object, newdata) {
    predict(object, data = newdata, type="response")$predictions[,"Deferred"]
  }

  # Not used currently as this doesn't perform preprocessing on newdata
  pfun_ksvm <- function(object, newdata) {
    predict(object, newdata = newdata, type="probabilities")[,2]
  }

  # This uses decision values instead of probabilities
  pfun_ksvm_decision <- function(object, newdata) {
    my_predict(object, newdata = newdata, type="decision")[,1]
  }
  
  # This is Caret's wrapper model
  pfun_train <- function(object, newdata) {
    predict(object, newdata = newdata, type="prob")[,2]
  }
  
  
  if ("randomForest" %in% class(model)) {
    pfun <- pfun_randomForest
  } else if ("ksvm" %in% class(model)) {
    pfun <- pfun_ksvm
  } else if ("ranger" %in% class(model)) {
    pfun <- pfun_ranger
  } else if ("train" %in% class(model)) {
    if ("ksvm" %in% class(model$finalModel)) {
      pfun <- pfun_ksvm_decision
    } else pfun <- pfun_train
  } else if ("stanfit" %in% class(model)) {
    pfun <- pfun_lmm
  }
  
  #print(pfun)
  
  result_code <- tryCatch(
    error = function(cnd) {
      msg <- paste("\nComputation of shap values failed:", cnd$message, 
                   sep="\n")
      warning(msg)
      NULL
    },
    {
      rlang::with_options(lifecycle_verbosity = "quiet", {  # Prevent the deprecation message caused by the explain function
        if ("lm" %in% class(model)) {
          shap <- fastshap::explain(model,   # This is for the baseline logistic regression
                                    feature_names = "previous_Hb",
                                    newdata = as.data.frame(validate2 %>% select(previous_Hb)),
                                    exact = TRUE)
        } else if ("stanfit" %in% class(model)) {
          shap <- fastshap::explain(model, 
                                    X = as.data.frame(validate2), 
                                    #newdata = as.data.frame(validate), 
                                    feature_names = setdiff(colnames(validate2), "donb"),
                                    pred_wrapper = pfun, 
                                    nsim = nsim)
        } else {
          shap <- fastshap::explain(model, 
                                    X = as.data.frame(validate2),
                                    pred_wrapper = pfun, 
                                    nsim = nsim)
        }
      })
      shap <- as_tibble(shap)  # This drops the class "explain"
    }
    
  )
  if (is.null(result_code)) {
    return(NULL)
  }
  
  if (any(shap %>% mutate(c=if_any(everything(), is.na)) %>% pull(c)) || nrow(shap) == 2*n) {
    warning("Predict function failed in compute_shap_values_fastshap. You could try rerunning with a different seed")
    return(NULL)
  }
  n <- nrow(validate2)
  attributions <- shap %>% mutate(id=1:n)
  attributions <- pivot_longer(attributions, cols=!id) %>%
    select(Variable=name, id, attribution=value)
  features <- validate2 %>% mutate(id=1:n)
  if ("sex" %in% colnames(validate2)) {
    features <- features %>% mutate(sex = sex=="female")
  }
  features <- features %>% pivot_longer(cols=!id) %>%
    select(Variable=name, id, value=value)
  res <- inner_join(attributions, features, by=c("Variable", "id")) %>%
    group_by(Variable) %>%
    mutate(value=scale(value)[,1]) %>%
    ungroup()
  
  res <- res %>% select(-id) %>% slice_sample(prop = 1.0)  # permute the rows
  
  return(res)
}
```



# Functions for calculating and plotting mean absolute attribution

```{r}
calc_mean_abs_attr <- function(df, seed) {
  df <- df %>%
    filter(Variable != 'sex') %>%
    mutate(abs_attr = abs(attribution)) %>%
    group_by(Variable) %>%
    summarise(mean_abs_attr = mean(abs_attr)) %>%
    ungroup() %>%
    mutate(seed = seed,
           Variable = as.factor(Variable),
           Variable = fct_relevel(Variable, 'previous_Hb', 'recent_deferrals',
                                  'Hb_first', 'days_to_previous_fb', 'recent_donations',
                                  'hour', 'previous_Hb_def', 'consecutive_deferrals',
                                  'warm_season', 'age'))
  return(list(df))
}

plot_maa_seeds <- function(df, title, full) {
  grob <- grobTree(textGrob('Cross shows value for full test set', x=0.6, y=0.05,
                   hjust=0, gp=gpar(col="black", fontsize=10, fontface="italic")))
  ggplot() +
    geom_point(data=df, aes(x=mean_abs_attr, y=Variable, color=as.character(seed))) +
    scale_colour_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c',
                                 '#fb9a99','#e31a1c','#fdbf6f','#ff7f00',
                                 '#cab2d6','#6a3d9a')) +
    scale_y_discrete(limits=rev) +
    ggtitle(paste('Mean absolute attribution in different seeds,', title)) +
    geom_point(data=full, aes(x=mean_abs_attr, y=Variable), shape=4, size=4) +
    theme_bw() + 
    theme(legend.position='none') +
    annotation_custom(grob)
}
```



# Loading datafiles

```{r}
path <- '../../results/multiple dutch 10k runs v2/'

folders <- list.dirs(path, full.names=FALSE, recursive=FALSE)
folders <- folders[folders != '']

folder <- folders[1] # We only need one run for this

svm_female <- load_model_validate(path, folder, 'svm', 'female')
svm_male <- load_model_validate(path, folder, 'svm', 'male')
rf_female <- load_model_validate(path, folder, 'rf', 'female')
rf_male <- load_model_validate(path, folder, 'rf', 'male')

seeds <- c(6652, 2070, 8719, 8473, 1966, 6247, 9593, 1607, 1411, 4375)
```

# Calculating SHAP values on the full validation set (n=2000)
These are then plotted alongside the SHAP values of the different subsamples.

```{r}
seeds <- c(6652, 2070, 8719, 8473, 1966, 6247, 9593, 1607, 1411, 4375)

tic()
res_svm_female_full <- compute_shap_values_fastshap(svm_female$model, 
                                                    svm_female$validate,
                                                    n=2000,
                                                    seed=123)
toc()
tic()
res_svm_male_full <- compute_shap_values_fastshap(svm_male$model, 
                                                  svm_male$validate,
                                                  n=2000,
                                                  seed=123)
toc()
tic()
res_rf_female_full <- compute_shap_values_fastshap(rf_female$model, 
                                                   rf_female$validate,
                                                   n=2000,
                                                   seed=123)
toc()
tic()
res_rf_male_full <- compute_shap_values_fastshap(rf_male$model, 
                                                 rf_male$validate,
                                                 n=2000,
                                                 seed=123)
toc()

saveRDS(res_svm_female_full, paste0(path, 'SHAP_res_svm_female_full.RDS'))
saveRDS(res_svm_male_full, paste0(path, 'SHAP_res_svm_male_full.RDS'))
saveRDS(res_rf_female_full, paste0(path, 'SHAP_res_rf_female_full.RDS'))
saveRDS(res_rf_male_full, paste0(path, 'SHAP_res_rf_male_full.RDS'))
```






# Compute SHAP values on same data/model, but using different seeds, n=100

```{r}
seeds <- c(6652, 2070, 8719, 8473, 1966, 6247, 9593, 1607, 1411, 4375)

# This could probably be nicer using apply...

res_svm_female <- list()
for (i in 1:length(seeds)) {
  tic()
  res <- compute_shap_values_fastshap(svm_female$model, 
                                      svm_female$validate, 
                                      n=100, 
                                      seed=seeds[i])
  res_svm_female[i] <- list(res)
  toc()
}

res_svm_male <- list()
for (i in 1:length(seeds)) {
  tic()
  res <- compute_shap_values_fastshap(svm_male$model, 
                                      svm_male$validate, 
                                      n=100, 
                                      seed=seeds[i])
  res_svm_male[i] <- list(res)
  toc()
}

res_rf_female <- list()
for (i in 1:length(seeds)) {
  tic()
  res <- compute_shap_values_fastshap(rf_female$model, 
                                      rf_female$validate, 
                                      n=100, 
                                      seed=seeds[i])
  res_rf_female[i] <- list(res)
  toc()
}

res_rf_male <- list()
for (i in 1:length(seeds)) {
  tic()
  res <- compute_shap_values_fastshap(rf_male$model, 
                                      rf_male$validate, 
                                      n=100, 
                                      seed=seeds[i])
  res_rf_male[i] <- list(res)
  toc()
}

saveRDS(res_svm_female, paste0(path, 'SHAP_res_svm_female_100.RDS'))
saveRDS(res_svm_male, paste0(path, 'SHAP_res_svm_male_100.RDS'))
saveRDS(res_rf_female, paste0(path, 'SHAP_res_rf_female_100.RDS'))
saveRDS(res_rf_male, paste0(path, 'SHAP_res_rf_male_100.RDS'))

```

# Load SHAP values for full validation set

```{r}
res_svm_female_full <- readRDS(paste0(path, 'SHAP_res_svm_female_full.RDS'))
res_svm_male_full <- readRDS(paste0(path, 'SHAP_res_svm_male_full.RDS'))
res_rf_female_full <- readRDS(paste0(path, 'SHAP_res_rf_female_full.RDS'))
res_rf_male_full <- readRDS(paste0(path, 'SHAP_res_rf_male_full.RDS'))

maa_svm_female_full <- calc_mean_abs_attr(res_svm_female_full, 0)[[1]]
maa_svm_male_full <- calc_mean_abs_attr(res_svm_male_full, 0)[[1]]
maa_rf_female_full <- calc_mean_abs_attr(res_rf_female_full, 0)[[1]]
maa_rf_male_full <- calc_mean_abs_attr(res_rf_male_full, 0)[[1]]
```


# Comparing SHAP values different seeds for n=100

```{r}
# Load SHAP values for n=100
res_svm_female <- readRDS(paste0(path, 'SHAP_res_svm_female_100.RDS'))
res_svm_male <- readRDS(paste0(path, 'SHAP_res_svm_male_100.RDS'))
res_rf_female <- readRDS(paste0(path, 'SHAP_res_rf_female_100.RDS'))
res_rf_male <- readRDS(paste0(path, 'SHAP_res_rf_male_100.RDS'))

maa_svm_female <- bind_rows(mapply(calc_mean_abs_attr, res_svm_female, seeds, SIMPLIFY=TRUE))
maa_svm_male <- bind_rows(mapply(calc_mean_abs_attr, res_svm_male, seeds, SIMPLIFY=TRUE))
maa_rf_female <- bind_rows(mapply(calc_mean_abs_attr, res_rf_female, seeds, SIMPLIFY=TRUE))
maa_rf_male <- bind_rows(mapply(calc_mean_abs_attr, res_rf_male, seeds, SIMPLIFY=TRUE))

plot_maa_seeds(maa_svm_female, 'SVM female, n=100', maa_svm_female_full)
ggsave('../effect_seed_shap_plots/maa_svm_female_n100.png')

plot_maa_seeds(maa_svm_male, 'SVM male, n=100', maa_svm_male_full)
ggsave('../effect_seed_shap_plots/maa_svm_male_n100.png')

plot_maa_seeds(maa_rf_female, 'RF female, n=100', maa_rf_female_full)
ggsave('../effect_seed_shap_plots/maa_rf_female_n100.png')

plot_maa_seeds(maa_rf_male, 'RF male, n=100', maa_rf_male_full)
ggsave('../effect_seed_shap_plots/maa_rf_male_n100.png')
```


# Calculating SHAP values for n=1000 for different seeds

```{r}
seeds <- c(6652, 2070, 8719, 8473, 1966, 6247, 9593, 1607, 1411, 4375)

res_svm_female_1000 <- list()
for (i in 1:length(seeds)) {
  tic()
  res <- compute_shap_values_fastshap(svm_female$model, 
                                      svm_female$validate, 
                                      n=1000, 
                                      seed=seeds[i])
  res_svm_female_1000[i] <- list(res)
  toc()
}

res_svm_male_1000 <- list()
for (i in 1:length(seeds)) {
  tic()
  res <- compute_shap_values_fastshap(svm_male$model, 
                                      svm_male$validate, 
                                      n=1000, 
                                      seed=seeds[i])
  res_svm_male_1000[i] <- list(res)
  toc()
}

res_rf_female_1000 <- list()
for (i in 1:length(seeds)) {
  tic()
  res <- compute_shap_values_fastshap(rf_female$model, 
                                      rf_female$validate, 
                                      n=1000, 
                                      seed=seeds[i])
  res_rf_female_1000[i] <- list(res)
  toc()
}

res_rf_male_1000 <- list()
for (i in 1:length(seeds)) {
  tic()
  res <- compute_shap_values_fastshap(rf_male$model, 
                                      rf_male$validate, 
                                      n=1000, 
                                      seed=seeds[i])
  res_rf_male_1000[i] <- list(res)
  toc()
}

saveRDS(res_svm_female_1000, paste0(path, 'SHAP_res_svm_female_1000.RDS'))
saveRDS(res_svm_male_1000, paste0(path, 'SHAP_res_svm_male_1000.RDS'))
saveRDS(res_rf_female_1000, paste0(path, 'SHAP_res_rf_female_1000.RDS'))
saveRDS(res_rf_male_1000, paste0(path, 'SHAP_res_rf_male_1000.RDS'))
```

# Comparing SHAP values for n=1000

```{r}
res_svm_female_1000 <- readRDS(paste0(path, 'SHAP_res_svm_female_1000.RDS'))
res_svm_male_1000 <- readRDS(paste0(path, 'SHAP_res_svm_male_1000.RDS'))
res_rf_female_1000 <- readRDS(paste0(path, 'SHAP_res_rf_female_1000.RDS'))
res_rf_male_1000 <- readRDS(paste0(path, 'SHAP_res_rf_male_1000.RDS'))

maa_svm_female <- bind_rows(mapply(calc_mean_abs_attr, res_svm_female_1000, seeds, SIMPLIFY=TRUE))
maa_svm_male <- bind_rows(mapply(calc_mean_abs_attr, res_svm_male_1000, seeds, SIMPLIFY=TRUE))
maa_rf_female <- bind_rows(mapply(calc_mean_abs_attr, res_rf_female_1000, seeds, SIMPLIFY=TRUE))
maa_rf_male <- bind_rows(mapply(calc_mean_abs_attr, res_rf_male_1000, seeds, SIMPLIFY=TRUE))

plot_maa_seeds(maa_svm_female, 'SVM female n=1000', maa_svm_female_full)
ggsave('../effect_seed_shap_plots/maa_svm_female_n1000.png')

plot_maa_seeds(maa_svm_male, 'SVM male n=1000', maa_svm_male_full)
ggsave('../effect_seed_shap_plots/maa_svm_male_n1000.png')

plot_maa_seeds(maa_rf_female, 'RF female n=1000', maa_rf_female_full)
ggsave('../effect_seed_shap_plots/maa_rf_female_n1000.png')

plot_maa_seeds(maa_rf_male, 'RF male n=1000', maa_rf_male_full)
ggsave('../effect_seed_shap_plots/maa_rf_male_n1000.png')
```










